Supplementary Information part 2: Testing the improved data sets

### 2022-08-11 paper/Forest_and_Water/SupplementaryMaterialPart2.Rmd
further updates paper 11082022

---
title: 'Supplementary Information part 2: Testing the improved data sets'
author: "Willem Vervoort/Eliana Nervi/Jimena Alonso"
date: "`r Sys.Date()`"
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output:   
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: forestandwater.bib
editor_options: 
  chunk_output_type: inline
header-includes:
  - \usepackage{setspace}
  - \usepackage{color}
abstract: |
  This supplementary material file compares whether the inclusion of additional catchments generates fundamentally different results as the original (but improved) data. Single variable regressions on the smaller (original) dataset are compared with the extended data set.
---

@@

Zhang_small <- read_csv("../../data/SmallCatchments_T2.csv") %>%
  mutate(table = "Small catchments")
Zhang_large <- read_csv("../../data/LargeCatchments_T1.csv") %>%
  mutate(table = "Large catchments")
new_data <- read_csv("../../data/NewCatchments_T3.csv") %>%
  mutate(table = "new catchments")

@@

### Implementing the changes to the overall data

@@

@@

$S_f = \frac{\Delta Q_f}{\Delta F}$

```{r, echo = T}
All_data <- All_data %>%
  filter(is.na(DeltaF_perc) == F) %>%
  mutate(S_f = DeltaQf_perc/DeltaF_perc)
```

@@

1. The change in flow as a function of change in forest cover

First the results for the original (improved) dataset

```{r, echo = F, message = F}
png("Fig2Zhang.png", width = 2500, height = 1800, res = 300)
All_data %>%
  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(DeltaF_perc,DeltaQf_perc, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Change in forest cover %") + ylab("Change in flow (Qf) %")
dev.off()

# and for all data sets
png("Fig2Zhang_all.png", width = 2500, height = 1800, res = 300)
All_data %>%
#  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(DeltaF_perc,DeltaQf_perc, colour = forest_sign, shape = table)) + 
  geom_point() +  geom_smooth(method = "lm") +
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Change in forest cover %") + ylab("Change in flow (Qf) %")
dev.off()

@@

include_graphics("Fig2Zhang.png")
```
And then the results for all the catchments including the new "added" catchments.

@@

include_graphics("Fig2Zhang_all.png")
```

>>>>

---
title: 'Supplementary Information part 2: Testing the improved data sets'
author:
  - name: R. Willem Vervoort
    email: willem.vervoort@sydney.edu.au
    footnote: 1
  - name: Eliana Nervi
    email: eliananervif@gmail.com
  - name: Jimena Alonso
    email: jalonso@fing.edu.uy
footnote:
  - code: 1
    text: "Corresponding Author"    
date: "`r Sys.Date()`"
linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output:   
  bookdown::pdf_book:
    base_format: rticles::elsevier_article
bibliography: forestandwater.bib
editor_options: 
  chunk_output_type: inline
header-includes:
  - \usepackage{setspace}
  - \usepackage{color}
abstract: |
  This supplementary material file compares whether the inclusion of additional catchments generates fundamentally different results as the original (but improved) data. Single variable regressions on the smaller (original) dataset are compared with the extended data set.
---

@@

Zhang_small <- read_csv("../../data/SmallCatchments_T2.csv") %>%
  mutate(table = "Small catchments")
Zhang_large <- read_csv("../../data/LargeCatchments_T1.csv") %>%
  mutate(table = "Large catchments")
new_data <- read_csv("../../data/NewCatchments_T3.csv") %>%
  mutate(table = ifelse(`Area(km2)` > 1000, "Large catchments", "Small catchments"))

@@

## Implementing the changes to the overall data

@@

## Approach and analyses

@@

$S_f = |\frac{\Delta Q_f}{\Delta F}|$

```{r, echo = T}
All_data <- All_data %>%
  filter(is.na(DeltaF_perc) == F) %>%
  mutate(S_f = abs(DeltaQf_perc/DeltaF_perc))
```

@@

## The change in stream flow as a function of change in forest cover

Figure \@ref(fig:Fig2Zhang) highlights that the overall relationship in the updated dataset is the same as in @zhang2017. This means that while the modifications have cleaned up the transcription errors in the data, they have not fundamentally changed the conclusions in the original paper.

The next figure (Figure \@ref(fig:Fig2Zhangnew)) is the same analysis, but this includes the new data that we identified in papers. Again, this figure highlights that the new datasets have not fundamentally changed the relationships found in @zhang2017.

```{r, echo = F, message = F, warning = F}
png("Fig2Zhang.png", width = 2500, height = 1800, res = 300)
All_data %>%
  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(DeltaF_perc,DeltaQf_perc, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Change in forest cover %") + ylab("Change in flow (Qf) %")
dev.off()

# and for all data sets
png("Fig2Zhang_all.png", width = 2500, height = 1800, res = 300)
All_data %>%
#  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(DeltaF_perc,DeltaQf_perc, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Change in forest cover %") + ylab("Change in flow (Qf) %")
dev.off()

@@

include_graphics("Fig2Zhang.png")
```

@@

include_graphics("Fig2Zhang_all.png")
```
## The relationship between the area of the catchment and the sensitivity of streamflow to the change in forest cover.

This analysis replicates Figure 3 in @zhang2017, which investigates for large and small catchments the sensitivity to runoff change from change in forest cover as a function of area. Note that in the original figure, the x-axis is on a log scale. In the original paper, the analysis is presented for all catchments as well as for large and small catchments. Here we only analyse the small and large catchments.

We can see from Figure \@ref(fig:Fig3Zhang) that again the updated database for the original dataset results in little change in the relationships for both large and small catchments. However, when the additional new catchments are added to the database (Figure \@ref(fig:Fig3Zhangnew)), the relationships clearly change. In particular, for small catchments gaining forest cover, the sensitivity appears positively correlated with the logarithm of the size of the catchments.

```{r, echo = F, message = F, warning = F}
png("Fig3Zhang.png", width = 2500, height = 1800, res = 300)
All_data %>%
  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(Area_km2, S_f, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  scale_x_continuous(trans='log10') +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("log10(Area in km2)") + ylab("Sf (sensivity)")
dev.off()

# and for all data sets
png("Fig3Zhang_all.png", width = 2500, height = 1800, res = 300)
All_data %>%
#  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(Area_km2,S_f, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  scale_x_continuous(trans='log10') +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("log10(Area in km2)") + ylab("Sf (sensitivity")
dev.off()

```{r Fig3Zhang, echo =F, fig.cap="Changes in flow based on the catchments from the original data set", out.width="90%"}

include_graphics("Fig3Zhang.png")
```

```{r Fig3Zhangnew, echo =F, fig.cap="Changes in flow based on the catchments from the extended data set", out.width="90%"}

include_graphics("Fig3Zhang_all.png")
```
## The sensitivity to forest loss as a function of dryness

The final analysis that we retest here is the relationship in Figure 4 in the original @zhang2017 paper, which highlights the sensitivity to forest loss as a function of dryness. We are again showing just the for the small and large catchments, similar to the original paper.

Similar to earlier analyses in this document Figure \@ref(fig:Fig4Zhang) show that the updated database for the original dataset results in little change in the relationships for both large and small catchments. However, when the additional new catchments are added to the database (Figure \@ref(fig:Fig4Zhangnew)), the relationships clearly change. In particular, for small catchments both for forest gains and losses the relationship changes and appears stronger.

```{r, echo = F, message = F, warning = F}
png("Fig4Zhang.png", width = 2500, height = 1800, res = 300)
All_data %>%
  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(Dryness, S_f, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  #scale_x_continuous(trans='log10') +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Dryness (E0/P)") + ylab("Sf (sensivity)")
dev.off()

# and for all data sets
png("Fig4Zhang_all.png", width = 2500, height = 1800, res = 300)
All_data %>%
#  filter(dataset == "original Zhang et al data") %>%
  ggplot(aes(Dryness,S_f, colour = forest_sign)) + 
  geom_point() +  geom_smooth(method = "lm") +
  #scale_x_continuous(trans='log10') +
  facet_wrap(~table, ncol=1, scales = "free") + 
  theme_bw() + 
    theme(axis.title = element_text(size = rel(1.5)),
        axis.text = element_text(size = rel(1.2)),
        legend.text = element_text(size = rel(1.2)),
        legend.title = element_text(size = rel(1.2)),
        strip.text = element_text(size = rel(1.5))) + 
  # scale_colour_manual(name = "Forest cover direction", 
  #                     values = c("Decrease" = "red","Increase" = "blue")) +
  xlab("Dryness (E0/P)") + ylab("Sf (sensitivity")
dev.off()

```{r Fig4Zhang, echo =F, fig.cap="Changes in flow based on the catchments from the original data set", out.width="90%"}

include_graphics("Fig4Zhang.png")
```

```{r Fig4Zhangnew, echo =F, fig.cap="Changes in flow based on the catchments from the extended data set", out.width="90%"}

include_graphics("Fig4Zhang_all.png")
```

### 2022-08-11 paper/Forest_and_Water/SupplementaryMaterialPart2.tex
further updates paper 11082022

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Willem Vervoort/Eliana Nervi/Jimena Alonso},
            pdftitle={Supplementary Information part 2: Testing the improved data sets},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{Supplementary Information part 2: Testing the improved data sets}
    \author[]{%
  %
  }

      \cortext[cor1]{Corresponding author}

@@

\hypertarget{implementing-the-changes-to-the-overall-data}{%
\subsubsection{Implementing the changes to the overall data}\label{implementing-the-changes-to-the-overall-data}}

@@

@@

\(S_f = \frac{\Delta Q_f}{\Delta F}\)

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(DeltaF\_perc) }\SpecialCharTok{==}\NormalTok{ F) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{S\_f =}\NormalTok{ DeltaQf\_perc}\SpecialCharTok{/}\NormalTok{DeltaF\_perc)}
\end{Highlighting}
\end{Shaded}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  The change in flow as a function of change in forest cover
\end{enumerate}

First the results for the original (improved) dataset

@@

And then the results for all the catchments including the new ``added'' catchments.

@@

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Supplementary Information part 2: Testing the improved data sets},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

@@

  \title{Supplementary Information part 2: Testing the improved data sets}
    \author[]{R. Willem Vervoort%
  %
  \fnref{1}}
   \ead{willem.vervoort@sydney.edu.au} 
    \author[]{Eliana Nervi%
  %
  }
   \ead{eliananervif@gmail.com} 
    \author[]{Jimena Alonso%
  %
  }
   \ead{jalonso@fing.edu.uy} 
      \cortext[cor1]{Corresponding author}
    \fntext[1]{Corresponding Author}

@@

\hypertarget{implementing-the-changes-to-the-overall-data}{%
\subsection{Implementing the changes to the overall data}\label{implementing-the-changes-to-the-overall-data}}

@@

\hypertarget{approach-and-analyses}{%
\subsection{Approach and analyses}\label{approach-and-analyses}}

@@

\(S_f = |\frac{\Delta Q_f}{\Delta F}|\)

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{All\_data }\OtherTok{\textless{}{-}}\NormalTok{ All\_data }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{filter}\NormalTok{(}\FunctionTok{is.na}\NormalTok{(DeltaF\_perc) }\SpecialCharTok{==}\NormalTok{ F) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{S\_f =} \FunctionTok{abs}\NormalTok{(DeltaQf\_perc}\SpecialCharTok{/}\NormalTok{DeltaF\_perc))}
\end{Highlighting}
\end{Shaded}

@@

\hypertarget{the-change-in-stream-flow-as-a-function-of-change-in-forest-cover}{%
\subsection{The change in stream flow as a function of change in forest cover}\label{the-change-in-stream-flow-as-a-function-of-change-in-forest-cover}}

Figure \ref{fig:Fig2Zhang} highlights that the overall relationship in the updated dataset is the same as in \citet{zhang2017}. This means that while the modifications have cleaned up the transcription errors in the data, they have not fundamentally changed the conclusions in the original paper.

The next figure (Figure \ref{fig:Fig2Zhangnew}) is the same analysis, but this includes the new data that we identified in papers. Again, this figure highlights that the new datasets have not fundamentally changed the relationships found in \citet{zhang2017}.

@@

@@

\hypertarget{the-relationship-between-the-area-of-the-catchment-and-the-sensitivity-of-streamflow-to-the-change-in-forest-cover.}{%
\subsection{The relationship between the area of the catchment and the sensitivity of streamflow to the change in forest cover.}\label{the-relationship-between-the-area-of-the-catchment-and-the-sensitivity-of-streamflow-to-the-change-in-forest-cover.}}

This analysis replicates Figure 3 in \citet{zhang2017}, which investigates for large and small catchments the sensitivity to runoff change from change in forest cover as a function of area. Note that in the original figure, the x-axis is on a log scale. In the original paper, the analysis is presented for all catchments as well as for large and small catchments. Here we only analyse the small and large catchments.

We can see from Figure \ref{fig:Fig3Zhang} that again the updated database for the original dataset results in little change in the relationships for both large and small catchments. However, when the additional new catchments are added to the database (Figure \ref{fig:Fig3Zhangnew}), the relationships clearly change. In particular, for small catchments gaining forest cover, the sensitivity appears positively correlated with the logarithm of the size of the catchments.

\begin{figure}
\includegraphics[width=0.9\linewidth]{Fig3Zhang} \caption{Changes in flow based on the catchments from the original data set}\label{fig:Fig3Zhang}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth]{Fig3Zhang_all} \caption{Changes in flow based on the catchments from the extended data set}\label{fig:Fig3Zhangnew}
\end{figure}

\hypertarget{the-sensitivity-to-forest-loss-as-a-function-of-dryness}{%
\subsection{The sensitivity to forest loss as a function of dryness}\label{the-sensitivity-to-forest-loss-as-a-function-of-dryness}}

The final analysis that we retest here is the relationship in Figure 4 in the original \citet{zhang2017} paper, which highlights the sensitivity to forest loss as a function of dryness. We are again showing just the for the small and large catchments, similar to the original paper.

Similar to earlier analyses in this document Figure \ref{fig:Fig4Zhang} show that the updated database for the original dataset results in little change in the relationships for both large and small catchments. However, when the additional new catchments are added to the database (Figure \ref{fig:Fig4Zhangnew}), the relationships clearly change. In particular, for small catchments both for forest gains and losses the relationship changes and appears stronger.

\begin{figure}
\includegraphics[width=0.9\linewidth]{Fig4Zhang} \caption{Changes in flow based on the catchments from the original data set}\label{fig:Fig4Zhang}
\end{figure}

\begin{figure}
\includegraphics[width=0.9\linewidth]{Fig4Zhang_all} \caption{Changes in flow based on the catchments from the extended data set}\label{fig:Fig4Zhangnew}
\end{figure}
